# Experiment 2 {#exp2}

## Materials

\FloatBarrier

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

We excluded all participants whose answer did not resemble "I pay attention".

## Descriptives

```{r exp2-descriptives}
# by discipline
exp2 %>% 
  group_by(discipline, impressiveness) %>% 
  summarise(across(c(impressed, learn, competence, trust, consensus), 
                   list(mean = ~mean(., na.rm=TRUE)), 
                   .names = "{col}_{fn}")) %>%
  mutate_if(is.numeric, ~round(.x, digits = 2)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = "Descriptives") %>%
  kable_styling(full_width = FALSE) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```

## Results

Fig. \@ref(fig:exp2-plot) summarizes descriptive results.

(ref:exp2-plot) Overview of treatment effect on various outcomes

```{r exp2-plot, fig.cap="(ref:exp2-plot)"}
plot_data <- exp2 %>% 
  pivot_longer(
    c(competence, trust, impressed, learn),
    names_to = "outcome", 
               values_to = "value")

ggplot(plot_data, aes(x = impressiveness, y = value, fill = impressiveness, 
                 color = impressiveness)) +
  geom_half_violin (position = position_nudge(x = -.05), 
                    adjust=1.5, alpha = .4,
                    side = "l") +
  geom_dotplot(position = position_nudge(x = -.05),
               binaxis = 'y', stackdir = 'down', dotsize = 0.1, binwidth = 0.5) + 
  # colors 
  scale_color_viridis_d(option = "turbo", begin = 0.1, end = 0.9) +
  scale_fill_viridis_d(option = "turbo", begin = 0.1,, end = 0.9) +
  guides(color = "none", fill = "none") +
  # labels
  labs(x = "Impressiveness", y = "Likert-scale rating") +
  plot_theme +
  facet_wrap(~outcome)


```

The results of the models from the hypotheses can be found in table \@ref(tab:exp2-hypotheses-table).

```{r exp2-hypotheses-table}
# H1 a
H1a_model <- lmer(competence ~ impressiveness + (1 | id), exp2)

# H1 b
H1b_model <- lmer(competence ~ impressed + (1 | id), exp2)

# H2 a
H2a_model <- lmer(trust ~ impressiveness + (1 | id), exp2)

# H2 b
H2b_model <- lmer(trust ~ impressed + (1 | id), exp2)

modelsummary(list(
                  "H1a (Competence)" = H1a_model,
                  "H1b (Competence pooled)" = H1b_model,
                  "H2a (Trust)" = H2a_model,
                  "H2b (Trust pooled)" = H2b_model
                  ), 
             stars = TRUE, 
             output = "kableExtra") %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```

The results of the model from RQ1 and of the models from RQ2 (interaction between impressiveness and consensus on competence/trust) can be found in table \@ref(tab:exp2-consensus-table).

```{r exp2-consensus-table}
# RQ1
RQ1_model <- lmer(learn ~ impressiveness + (1 | id), exp2)

# RQ2
# for H1
RQ2_H1a_model <- lmer(trust ~ impressiveness*consensus + (1 | id), exp2)
RQ2_H1b_model <- lmer(trust ~ impressed*consensus + (1 | id), exp2)
# for H2
RQ2_H2a_model <- lmer(competence ~ impressiveness*consensus + (1 | id), exp2)
RQ2_H2b_model <- lmer(competence~ impressed*consensus + (1 | id), exp2)

modelsummary(list("RQ 1" = RQ1_model,
                  "H1a x Consensus" = RQ2_H1a_model,
                  "H1b x Consensus" = RQ2_H1b_model,
                  "H2a x Consensus" = RQ2_H2a_model,
                  "H2b x Consensus" = RQ2_H2b_model
), 
stars = TRUE, 
output = "kableExtra") %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```
