# Experiment 1b {#exp1b}

This experiment is not reported in the main article, but similar in Design to Experiment 1 reported in the article, which is why we refer to it as Experiment 1b. Please note that on the OSF, all files related to this Experiment 1 are under "experiment_2" and files related to Experiment 1b under "experiment_1", as this was the order in which experiments were run. 

## Differences with Experiment 1 from the main article

\FloatBarrier

Experiment 1 reported in the main article is essentially a slightly improved version of Experiment 1b. Although we confirmed all our hypotheses, there were a couple of issues with Experiment 1b that make us feel that a replication was warranted: First, we made a technical mistake: In the entomology condition, our manipulation check question read: “How impressive do you think the findings of the archeologists described in the text are?”, whereas it should have read “… the findings of entomologists…”. Two participants brought this issue to our attention. While it should not have affected Hypotheses 1a and 2a, it might have distorted results on Hypotheses 1b and 2b which rely on the faulty question item. We corrected this mistake for Experiment 1. 

Second, in Experiment 1b, participants who read the archeology vignettes--contrary to the findings of a pilot study--did not rate the impressive version as more impressive than the basic version. We therefore used slightly modified vignettes for archeology for Experiment 1, hoping to make the two conditions more distinguishable. We also shortened the text in both conditions to be more similar in length to (i) each other and (ii) to the entomology vignettes. Table \@ref(tab:exp1b-stimuli) shows the vignettes used in Experiment 1b. Table \@ref(tab:exp1-exp1b-differences) highlights the differences in the archeology vignettes between Experiment 1 and 1b. Third, for a minor issue, one participant remarked that we should have spelled “archaeology” instead of “archeology”. To avoid potential confusion, we changed the spelling accordingly in Experiment 1.

By contrast to Experiment 1, we ran Experiment 1b on a sample of UK participants, using the opportunity of replicating the main results on a sample from a different population.

The vignettes about entomology remained the same as in Experiment 1 (see Table \@ref(tab:exp1-stimuli)).

The main goal of experiment one to see if exposure to impressive science content could enhance people's trust in scientists. We had the following hypotheses:

**H1a: Participants will perceive scientists as more competent than they did before after having read an impressive text about their discipline's findings, compared to when reading a basic text.**

**H1b: Across all conditions, participants who are more impressed by the text about a discipline will also tend to perceive the scientists of the discipline as more competent.**

**H2a: Participants will trust a discipline more than they did before, after reading an impressive text about the discipline's findings, compared to when reading a basic text.**

**H2b: Across all conditions, participants who are more impressed by the text about a discipline will also tend to trust the scientists of the discipline more.**

We had the following research questions:

**RQ1: Do participants perceive that they learn more from the texts in the impressive condition, compared to the basic condition?**

**RQ2: Do perceptions of consensus interact with the relationships proposed in the hypotheses, such that greater perceived consensus is associated with a more positive relationship between impressiveness and trust/competence?**

```{r exp1b}
# Analyze data of experiments and store results

# Experiment 1

# read the cleaned version data set
exp1b <- read_csv("experiment_1/data/cleaned.csv")

# Hypotheses

# H1 a
H1a_model <- lmer(competence ~ impressiveness + (1 | id), exp1b)

# H1 b
H1b_model <- lmer(competence ~ impressed + (1 | id), exp1b)

# H2 a
H2a_model <- lmer(trust ~ impressiveness + (1 | id), exp1b)

# H2 b
H2b_model <- lmer(trust ~ impressed + (1 | id), exp1b)

# Research questions

# RQ1
RQ1_model <- lmer(learn ~ impressiveness + (1 | id), exp1b)

# RQ2
# for H1
RQ2_H1a_model <- lmer(trust ~ impressiveness*consensus + (1 | id), exp1b)
RQ2_H1b_model <- lmer(trust ~ impressed*consensus + (1 | id), exp1b)
# for H2
RQ2_H2a_model <- lmer(competence ~ impressiveness*consensus + (1 | id), exp1b)
RQ2_H2b_model <- lmer(competence~ impressed*consensus + (1 | id), exp1b)

# Manipulation check
manipulation_check_model <- lmer(impressed ~ impressiveness + (1 | id), exp1b) 

# by discipline
manipulation_check_archeo <- lm(impressed ~ impressiveness, 
                                  exp1b %>% filter(discipline == "archeo")) 

manipulation_check_entom <- lm(impressed ~ impressiveness, 
                                  exp1b %>% filter(discipline == "entom")) 

# extract descriptives for inline reporting
exp1b_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp1b$id),
  gender = exp1b %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp1b %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # Means
  means = exp1b %>% 
    group_by(impressiveness) %>% 
    summarize(across(c(competence, trust, impressed, learn), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    mutate_if(is.numeric, round, digits = 2) %>% 
    split(.$impressiveness),
  means_by_discipline = exp1b %>% 
    group_by(impressiveness, discipline) %>% 
    summarize(across(c(competence, trust, impressed, learn), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    mutate_if(is.numeric, round, digits = 2) %>% 
    super_split(impressiveness, discipline),
  # H1
  H1a = text_ready(H1a_model),
  H1b = text_ready(H1b_model),
  # H2
  H2a = text_ready(H2a_model),
  H2b = text_ready(H2b_model),
  # RQ1
  RQ1 = text_ready(RQ1_model),
  # RQ2
  RQ2_H1a = text_ready(RQ2_H1a_model), 
  RQ2_H1b = text_ready(RQ2_H1b_model), 
  RQ2_H2a = text_ready(RQ2_H2a_model), 
  RQ2_H2b = text_ready(RQ2_H2b_model),
  # Manipulation check
  manipulation_check = text_ready(manipulation_check_model),
  manipulation_check_archeo = text_ready(manipulation_check_archeo),
  manipulation_check_entom = text_ready(manipulation_check_entom)
)
```

## Methods

### Participants

We recruited 100 participants from the US via Prolific. Two participants failed our attention check, resulting in a final sample of `r exp1b_descriptives$n_subj` participants (`r exp1b_descriptives$gender$female$n` female, `r exp1b_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp1b_descriptives$age$mean`, $age_\text{sd}$: `r exp1b_descriptives$age$sd`, $age_\text{median}$: `r exp1b_descriptives$age$median`).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "You’re going to read two texts about scientific discoveries, in two fields: archeology (the study of human activity through the recovery and analysis of material culture), and entomology (the study of insects). We ask you to read the texts, and then answer a few questions about them." Next, participants read vignettes about scientific findings in the disciplines of entomology and archaeology. We manipulated the impressiveness of the texts, so that there was one "basic" and one "impressive" version for each of the disciplines (see Table \@ref(tab:exp1b-stimuli)). We varied impressiveness within participants, but between disciplines: each participant saw an impressive version for one discipline, and a basic version for the other discipline. After reading each text, participants were asked a series of questions. We asked participants how much they thought they learned from reading the text ("How much do you feel you’ve learnt about human history by reading this text?" [1 - Nothing, 2 - A bit, 3 - Some, 4 - Quite a bit, 5 - A lot]) and how impressive they found it (“How impressive do you think the findings of the archaeologists described in the text are?” [1 - Not very impressive, 2 - A bit impressive, 3 - Quite impressive, 4 - Very impressive, 5 - Extremely impressive]). We also asked them whether reading the text changed their impression of the competence of the scientists of the respective discipline ("Would you agree that reading this text has made you think of archaeologists as more competent than you thought before?" [1 - Strongly disagree, 2 - Disagree, 3 - Neither agree nor disagree, 4 - Agree, 5 - Strongly agree]) and their trust in the respective discipline ("Having read this text, would you agree that you trust the discipline of archaeology more than you did before?" [1 - Strongly disagree, 2 - Disagree, 3 - Neither agree nor disagree, 4 - Agree, 5 - Strongly agree]). Finally, we also asked about the perceived consensus ("To which extent do you think the findings from the short text you just read reflect a minority or a majority opinion among archaeologists?" [1 - Small minority, 2 - Minority, 3 - About half, 4 - Majority, 5 - Large majority]).

### Materials

#### Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

We excluded all participants whose answer did not resemble "I pay attention".

#### Vignettes

Table \@ref(tab:exp1b-stimuli) shows the vignettes used in Experiment 1b. Table \@ref(tab:exp1-exp1b-differences) highlights the differences in the archeology vignettes between Experiment 1 and 1b.

```{r exp1b-stimuli}
# Create a matrix of image file paths as Markdown-formatted strings
table <- data.frame(discipline = c("Archeology", 
                                   "Entomology"),
                    impressive = c("Archeologists, scientists who study human history and prehistory, are able to tell, from their bones, whether someone was male or female, how old they were, and whether they suffered from a wide range of diseases. Archeologists can now even tell at what age someone, dead for tens of thousands of years, stopped drinking their mother’s milk, from the composition of their teeth. 

Looking for patterns, archeologists can also understand how far our ancestors traveled, or how often they fought each other.

Archeologists can learn about the language that our ancestors or cousins might have had. For instance, the nerve that is used to control breathing is larger in humans than in apes, plausibly because we need more fine-grained control of our breathing in order to speak. As a result, the canal containing that nerve is larger in humans than in apes – and it is also enlarged in Neanderthals. 

We can also tell, from an analysis of the tools they made, that, like modern humans, most Neanderthals were right-handed. It’s thought that handedness is related to the evolution of language, another piece of evidence suggesting that Neanderthals likely possessed a form of language. ", 
"Entomologists are the scientists who study insects. Some of them have specialized in understanding how insects perceive the world around them, and they have uncovered remarkable abilities. 

Entomologists interested in how flies’ visual perception works have used special displays to present images for much less than the blink of an eye, electrodes to record how individual cells in the flies’ brain react, and ultra-precise electron microscopy to examine their eyes. Thanks to these techniques, they have shown that some flies can perceive images that are displayed for just three milliseconds (a thousandth of a second) – about ten times shorter than a single movie frame (of which there are 24 per second). 

Entomologists who study the hair of crickets have shown that these microscopic hairs, which can be found on antenna-like organs attached to the crickets’ rear, are maybe the most sensitive organs in the animal kingdom. The researchers used extremely precise techniques to measure how the hair reacts to stimuli, such as laser-Doppler velocimetry, a technique capable of detecting the most minute of movements. They were able to show that the hair could react to changes in the motion of the air that had less energy than one particle of light, a single photon."),
                    basic = c("Archaeology is the science that studies human history and prehistory based on the analysis of objects from the past such as human bones, engravings, constructions, and various objects, from nails to bits of pots. This task requires a great deal of carefulness, because objects from the past need to often be dug out from the ground and cleaned, without destroying them in the process. 

Archeologists have been able to shed light on human history in all continents, from ancient Egypt to the Incas in Peru or the Khmers in Cambodia. 

Archaeologists have made some startling discoveries, such as the amazing animal paintings in the Lascaux caves, which have been shown to be at least 30000 years old. On that basis, archeologists speculate on the artistic and religious lives of our ancestors. 

Archeologists have also been able to find remains of our more distant ancestors, showing that our species is just one among several that appeared and then went extinct, such as Neanderthals, Homo erectus, or Homo habilis. 

Archaeology relies on scientific methods of analysis such as carbon dating, which enables us to date objects, based on the type of carbon atoms they contain.", 
"Entomologists are scientists who investigate insects, typically having a background in biology. They study, for example, how a swarm of bees organizes, or how ants communicate with each other. 

They also study how different insects interact with each other and their environment, whether some species are in danger of going extinct, or whether others are invasive species that need to be controlled.

Sometimes entomologists study insects by observing them in the wild, sometimes they conduct controlled experiments in laboratories, to see for example how different environmental factors change the behavior of insects, or to track exactly the same insects over a longer period of time.

An entomologist often specializes in one type of insect in order to study it in depth. For example, an entomologist who specializes in ants is called a myrmecologist. "))

# Output the table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kable, column_spec, and kable_styling
  table_output <- kableExtra::kable(table,
                  col.names = c("", "Impressive", "Basic"),
                  caption = "Stimuli of Experiment 1b",
                  align = "l",
                  booktabs = TRUE,
                  longtable = TRUE,
                  #full_width = TRUE,
                  #format = "pandoc"
                  ) %>%
  kable_styling(latex_options = "repeat_header",
                font_size = 8) %>%
  column_spec(1, width = "5em") %>%
    column_spec(2, width = "25em") %>%
    column_spec(3, width = "25em")
} else {
  # For Word: Use flextable and rename columns
  table_output <- flextable::flextable(table |> 
                                         rename(`Impressive text` = impressive, 
                                                `Basic text` = basic, 
                                                `Discipline` = discipline)) %>%
    flextable::set_caption(caption = "Stimuli of Experiment 1b") %>%
    flextable::autofit() %>%
    flextable::width(j = 1, width = 3) %>%
    flextable::width(j = 2, width = 5) %>%
    flextable::width(j = 3, width = 5)
}

table_output

```

```{r exp1-exp1b-differences, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
table <- data.frame(impressiveness = c("basic", 
                                       "impressive"),
                    Experiment_1 = c("Archaeology is the science that studies human history and prehistory based on the analysis of objects from the past such as human bones, engravings, constructions, and various objects, from nails to bits of pots. This task requires a great deal of carefulness, because objects from the past need to often be dug out from the ground and cleaned, without destroying them in the process. 

Archeologists have been able to shed light on human history in all continents, from ancient Egypt to the Incas in Peru or the Khmers in Cambodia. 

Archaeologists have made some startling discoveries, such as the amazing animal paintings in the Lascaux caves, which have been shown to be at least 30000 years old. On that basis, archeologists speculate on the artistic and religious lives of our ancestors. 

Archeologists have also been able to find remains of our more distant ancestors, showing that our species is just one among several that appeared and then went extinct, such as Neanderthals, Homo erectus, or Homo habilis. 

Archaeology relies on scientific methods of analysis such as carbon dating, which enables us to date objects, based on the type of carbon atoms they contain.", 
"Archeologists, scientists who study human history and prehistory, are able to tell, from their bones, whether someone was male or female, how old they were, and whether they suffered from a wide range of diseases. Archeologists can now even tell at what age someone, dead for tens of thousands of years, stopped drinking their mother’s milk, from the composition of their teeth. 

Looking for patterns, archeologists can also understand how far our ancestors traveled, or how often they fought each other.

Archeologists can learn about the language that our ancestors or cousins might have had. For instance, the nerve that is used to control breathing is larger in humans than in apes, plausibly because we need more fine-grained control of our breathing in order to speak. As a result, the canal containing that nerve is larger in humans than in apes – and it is also enlarged in Neanderthals. 

We can also tell, from an analysis of the tools they made, that, like modern humans, most Neanderthals were right-handed. It’s thought that handedness is related to the evolution of language, another piece of evidence suggesting that Neanderthals likely possessed a form of language.
"),
Experiment_2 = c("Archaeology is the science that studies human history and prehistory based on the analysis of objects from the past such as human bones, engravings, constructions, and various objects, from nails to bits of pots. This task requires a great deal of carefulness, because objects from the past need to often be dug out from the ground and patiently cleaned, without destroying them in the process.

Archaeologists have been able to shed light on human history in all continents, from ancient Egypt to the Incas in Peru or the Khmers in Cambodia.

Archaeologists study the paintings made by our ancestors, such as those that can be found in Lascaux, a set of caves found in the south of France that have been decorated by people at least 30000 years ago.

Archaeologists have also found remains of our more distant ancestors, showing that our species is just one among several that appeared, and then either changed or went extinct, such as Neanderthals, Homo erectus, or Homo habilis.", 
"Archaeologists, scientists who study human history and prehistory, are able to tell, from their bones, whether someone was male or female, how old they were, and whether they suffered from a range of diseases. Archaeologists can now tell at what age someone, dead for tens of thousands of years, stopped drinking their mother’s milk, from the composition of their teeth.
Archaeologists learn about the language that our ancestors or cousins might have had. For instance, the nerve that is used to control breathing is larger in humans than in apes, plausibly because we need more fine-grained control of our breathing in order to speak. As a result, the canal containing that nerve is larger in humans than in apes – and it is also enlarged in Neanderthals.
Archaeologists can also tell, from an analysis of the tools they made, that most Neanderthals were right-handed. It’s thought that handedness is related to the evolution of language, another piece of evidence suggesting that Neanderthals likely possessed a form of language."))


# Output the table
if (knitr::is_latex_output() || knitr::is_html_output()) {
  # For LaTeX or HTML: Use kable, column_spec, and kable_styling
  table_output <- kableExtra::kable(table, 
                  col.names = c("", "Experiment 1b", "Experiment 1"),
                  caption = "Differences in stimuli for Experiment 1 and 1b",
                  align = "l", 
                  booktabs = TRUE,
                  longtable = TRUE,
                  #full_width = TRUE,
                  #format = "pandoc"
                  ) %>%
  kable_styling(latex_options = "repeat_header",
                font_size = 8) %>%
  column_spec(1, width = "5em") %>%
    column_spec(2, width = "25em") %>%
    column_spec(3, width = "25em")
} else {
  # For Word: Use flextable and rename columns
  table_output <- flextable::flextable(table |> 
                                         rename(`Impressive text` = impressiveness, 
                                                `Experiment 1b` = Experiment_1, 
                                                `Experiment 1` = Experiment_2)) %>%
    flextable::set_caption(caption = "Differences in stimuli for Experiment 1 and 1b") %>%
    flextable::autofit() %>%
    flextable::width(j = 1, width = 3) %>%
    flextable::width(j = 2, width = 5) %>%
    flextable::width(j = 3, width = 5)
}

table_output
```

## Results and discussion

For a manipulation check, we find that--in line with pilot studies--participants indeed perceive the impressive text as more impressive (mean = `r exp1b_descriptives$means$imp$impressed_mean`, sd = `r exp1b_descriptives$means$imp$impressed_sd`; $\hat{b}$ = `r exp1b_descriptives$manipulation_check$impressivenessimp$estimate` `r exp1b_descriptives$manipulation_check$impressivenessimp$ci`, p `r exp1b_descriptives$manipulation_check$impressivenessimp$p.value`) than the basic text (mean = `r exp1b_descriptives$means$basic$impressed_mean`, sd = `r exp1b_descriptives$means$basic$impressed_sd`). However, this overall effect is driven by the entomology vignettes ($\hat{b}$ = `r exp1b_descriptives$manipulation_check_entom$impressivenessimp$estimate` `r exp1b_descriptives$manipulation_check_entom$impressivenessimp$ci`, p `r exp1b_descriptives$manipulation_check_entom$impressivenessimp$p.value`; $\text{mean}_{\text{basic}}$ = `r exp1b_descriptives$means_by_discipline$basic$entom$impressed_mean`, $\text{sd}_{\text{basic}}$ = `r exp1b_descriptives$means_by_discipline$basic$entom$impressed_sd`, $\text{mean}_{\text{impressive}}$ = `r exp1b_descriptives$means_by_discipline$imp$entom$impressed_mean`, $\text{sd}_{\text{impressive}}$ = `r exp1b_descriptives$means_by_discipline$imp$entom$impressed_sd`). For the archaeology vignettes we do not find a difference ($\hat{b}$ = `r exp1b_descriptives$manipulation_check_archeo$impressivenessimp$estimate` `r exp1b_descriptives$manipulation_check_archeo$impressivenessimp$ci`, p `r exp1b_descriptives$manipulation_check_archeo$impressivenessimp$p.value`; $\text{mean}_{\text{basic}}$ = `r exp1b_descriptives$means_by_discipline$basic$archeo$impressed_mean`, $\text{sd}_{\text{basic}}$ = `r exp1b_descriptives$means_by_discipline$basic$archeo$impressed_sd`, $\text{mean}_{\text{impressive}}$ = `r exp1b_descriptives$means_by_discipline$imp$archeo$impressed_mean`, $\text{sd}_{\text{impressive}}$ = `r exp1b_descriptives$means_by_discipline$imp$archeo$impressed_sd`).

We find that participants perceive scientists as more competent after having read an impressive text (mean = `r exp1b_descriptives$means$imp$competence_mean`, sd = `r exp1b_descriptives$means$imp$competence_sd`; $\hat{b}_{\text{Competence}}$ = `r exp1b_descriptives$H1a$impressivenessimp$estimate` `r exp1b_descriptives$H1a$impressivenessimp$ci`, p `r exp1b_descriptives$H1a$impressivenessimp$p.value`) than after having read a basic one (mean = `r exp1b_descriptives$means$basic$competence_mean`, sd = `r exp1b_descriptives$means$basic$competence_sd`). Pooled across all conditions, participants impressiveness ratings are positively associated with competence ($\hat{b}$ = `r exp1b_descriptives$H1b$impressed$estimate` `r exp1b_descriptives$H1b$impressed$ci`, p `r exp1b_descriptives$H1b$impressed$p.value`).

Similar to competence, we find that participants trust a discipline more after having read an impressive text (mean = `r exp1b_descriptives$means$imp$trust_mean`, sd = `r exp1b_descriptives$means$imp$trust_sd`; $\hat{b}_{\text{trust}}$ = `r exp1b_descriptives$H2a$impressivenessimp$estimate` `r exp1b_descriptives$H2a$impressivenessimp$ci`, p `r exp1b_descriptives$H2a$impressivenessimp$p.value`) than after having read a basic one. Participants impressiveness ratings are positively associated with trust when pooling across all conditions ($\hat{b}$ = `r exp1b_descriptives$H2b$impressed$estimate` `r exp1b_descriptives$H2b$impressed$ci`, p `r exp1b_descriptives$H2b$impressed$p.value`).

For RQ1, we find that participants had the impression of having learned more after having read the impressive text (mean = `r exp1b_descriptives$means$imp$learn_mean`, sd = `r exp1b_descriptives$means$imp$learn_sd`; $\hat{b}$ = `r exp1b_descriptives$RQ1$impressivenessimp$estimate` `r exp1b_descriptives$RQ1$impressivenessimp$ci`, p `r exp1b_descriptives$RQ1$impressivenessimp$p.value`), than after having read the basic one (mean = `r exp1b_descriptives$means$basic$learn_mean`, sd = `r exp1b_descriptives$means$basic$learn_sd`).

Regarding RQ2, we do not find evidence that consensus modulates the effect of impressiveness hypothesized in H1a and H2a. In Appendix \@ref(exp1b), we provide regression tables and figures.

### Summary tables and plots

```{r exp1b-descriptives}
# by discipline
exp1b %>% 
  group_by(discipline, impressiveness) %>% 
  summarise(across(c(impressed, learn, competence, trust, consensus), 
                   list(mean = ~mean(., na.rm=TRUE)), 
                   .names = "{col}_{fn}")) %>%
  mutate_if(is.numeric, ~round(.x, digits = 2)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = "Descriptives") %>%
  kable_styling(full_width = FALSE) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```

Fig. \@ref(fig:exp1b-plot) summarizes descriptive results.

(ref:exp1b-plot) Overview of treatment effect on various outcomes.

```{r exp1b-plot, fig.cap="(ref:exp1b-plot)"}
plot_data <- exp1b %>% 
  pivot_longer(
    c(competence, trust, impressed, learn),
    names_to = "outcome", 
               values_to = "value")

ggplot(plot_data, aes(x = impressiveness, y = value, fill = impressiveness, 
                 color = impressiveness)) +
  geom_half_violin (position = position_nudge(x = -.05), 
                    adjust=1.5, alpha = .4,
                    side = "l") +
  geom_dotplot(position = position_nudge(x = -.05),
               binaxis = 'y', stackdir = 'down', dotsize = 0.1, binwidth = 0.5) + 
  # colors 
  scale_color_viridis_d(option = "turbo", begin = 0.1, end = 0.9) +
  scale_fill_viridis_d(option = "turbo", begin = 0.1,, end = 0.9) +
  guides(color = "none", fill = "none") +
  # labels
  labs(x = "Impressiveness", y = "Likert-scale rating") +
  plot_theme +
  facet_wrap(~outcome)


```

The results of the models from the hypotheses can be found in table \@ref(tab:exp1b-hypotheses-table).

```{r exp1b-hypotheses-table}
# H1 a
H1a_model <- lmer(competence ~ impressiveness + (1 | id), exp1b)

# H1 b
H1b_model <- lmer(competence ~ impressed + (1 | id), exp1b)

# H2 a
H2a_model <- lmer(trust ~ impressiveness + (1 | id), exp1b)

# H2 b
H2b_model <- lmer(trust ~ impressed + (1 | id), exp1b)


# Generate the table with the correct format
if (knitr::is_latex_output() || knitr::is_html_output()) {
  modelsummary(list(
    "H1a (Competence)" = H1a_model,
    "H1b (Competence pooled)" = H1b_model,
    "H2a (Trust)" = H2a_model,
    "H2b (Trust pooled)" = H2b_model
  ), 
  stars = TRUE, 
  caption = "The results of the models for the hypotheses",
  output = "kableExtra") %>%
    # make smaller to fit
    kable_styling(latex_options = "scale_down")
  
} else {
  
  modelsummary(list(
    "H1a (Competence)" = H1a_model,
    "H1b (Competence pooled)" = H1b_model,
    "H2a (Trust)" = H2a_model,
    "H2b (Trust pooled)" = H2b_model
  ), 
  stars = TRUE,
  caption = "The results of the models for the hypotheses",
  output = "flextable") %>%
    flextable::set_caption("The results of the models from the hypotheses") %>%
    flextable::autofit() %>%  # Ensure proper column width
    flextable::theme_booktabs()  # Use a professional-looking theme
  
}
```

The results of the model from RQ1 and of the models from RQ2 (interaction between impressiveness and consensus on competence/trust) can be found in table \@ref(tab:exp1b-consensus-table).

```{r exp1b-consensus-table}
# RQ1
RQ1_model <- lmer(learn ~ impressiveness + (1 | id), exp1b)

# RQ2
# for H1
RQ2_H1a_model <- lmer(trust ~ impressiveness*consensus + (1 | id), exp1b)
RQ2_H1b_model <- lmer(trust ~ impressed*consensus + (1 | id), exp1b)
# for H2
RQ2_H2a_model <- lmer(competence ~ impressiveness*consensus + (1 | id), exp1b)
RQ2_H2b_model <- lmer(competence~ impressed*consensus + (1 | id), exp1b)


# Generate the table with the correct format
if (knitr::is_latex_output() || knitr::is_html_output()) {
  modelsummary(list("RQ 1" = RQ1_model,
                    "H1a x Consensus" = RQ2_H1a_model,
                    "H1b x Consensus" = RQ2_H1b_model,
                    "H2a x Consensus" = RQ2_H2a_model,
                    "H2b x Consensus" = RQ2_H2b_model
  ), 
  stars = TRUE, 
  caption = "The results of the model from RQ1 and of the models from RQ2 (interaction between impressiveness and consensus on competence/trust)",
  output = "kableExtra") %>%
    # make smaller to fit
    kable_styling(latex_options = "scale_down")
  
} else {
  
  modelsummary(list("RQ 1" = RQ1_model,
                    "H1a x Consensus" = RQ2_H1a_model,
                    "H1b x Consensus" = RQ2_H1b_model,
                    "H2a x Consensus" = RQ2_H2a_model,
                    "H2b x Consensus" = RQ2_H2b_model
  ), 
  stars = TRUE, 
  caption = "The results of the model from RQ1 and of the models from RQ2 (interaction between impressiveness and consensus on competence/trust)",
  output = "flextable") %>%
    flextable::set_caption("The results of the model from RQ1 and of the models from RQ2 (interaction between impressiveness and consensus on competence/trust)") %>%
    flextable::autofit() %>%  # Ensure proper column width
    flextable::theme_booktabs()  # Use a professional-looking theme
  
}

```
