# Stability of trust in science in the US {#stability}

\FloatBarrier

Figure \@ref(fig:gss-trend) shows how, since the 1970s, on average `r top_intitutions_great_deal$consci$mean_share` of Americans say they have a great deal of confidence in the scientific community, the second highest score among `r n_institutions` institutions included in the US General Social Survey (GSS), just behind the military.

Figure \@ref(fig:gss-trend-confidence) shows that the trend of stability for the share of Americans having a "great deal of confidence" in the scientific community also holds for those who have "only some" or "hardly any" confidence.

(ref:gss-trend) *Top four trusted institutions in the US*. The plot shows the variation of confidence in the top four trusted institutions in the US over time. The data for this plot is from the cumulative file of the US General Social Survey (GSS).

```{r gss-trend, fig.cap="(ref:gss-trend)"}
# Reduce data to top 4 trusted institutions
plot_data <- gss_summary_great_deal |>
  filter(institution %in% top_trusted_institutions$institution)

plot_data |> 
  ggplot(aes(x = year, y = share, colour = institution_label)) +
  # Background: Other institutions with lower alpha
  geom_line(data = plot_data |> 
              filter(institution != "consci"),
            aes(color = institution_label), alpha = 0.2, size = 1) + 

  # Foreground: Highlight "consci" with full opacity
  geom_line(data = plot_data |> 
              filter(institution == "consci"),
            aes(color = institution_label), size = 1.2) +

  geom_point(data = plot_data |> 
               filter(institution == "consci"),
             aes(color = institution_label), size = 2) +

  # Apply Wes Anderson colors
  scale_color_manual(values = wes_palette("AsteroidCity1", n = 4, type = "discrete")) +  

  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +  
  scale_x_continuous(breaks = c(seq(min(gss_summary_great_deal$year), max(gss_summary_great_deal$year), by = 5), max(gss_summary_great_deal$year))) +  

  labs(x = "Year", y = "Share of Americans \n with 'A great deal confidence'", 
       caption = "Data source: GSS by NORC") +
  theme_minimal() +  
  theme(legend.position = "none") +

  # Institution name nudged to the right
  geom_text_repel(
    data = plot_data |> 
      group_by(institution_label) |> 
      slice_max(year, n = 1, with_ties = FALSE),  
    aes(label = institution_label, 
        alpha = ifelse(institution == "consci", 1, 0.6)),  
    hjust = 0,  
    nudge_x = 5,  
    size = 3,  
    direction = "both",
    fontface = "bold",  
    segment.color = "grey50"  
  ) +
  scale_alpha_identity()
```

```{r}
# Check evolution for confidence in scientific community for all response options
gss_summary_polarization <- gss |>
  drop_na(confidence) |> 
  group_by(year, institution, institution_label, confidence_factor) |>
  filter(institution_label == "Scientific community") |> 
  summarize(
    n_responses_in_category = n()
  ) |>
  group_by(year, institution, institution_label) |> 
  # remove years without measure
    mutate(
      n_responses_in_year = sum(n_responses_in_category),
      share = n_responses_in_category/n_responses_in_year,
      # remove 0s
      share = ifelse(share == 0, NA, share)) |> 
  # make confidence_factor a factor with levels
  mutate(confidence_factor = factor(confidence_factor, 
                                    levels = c("Hardly any confidence at all", 
                                               "Only some confidence", 
                                               "A great deal confidence"),
                                    ordered = TRUE)) |> 
drop_na(share)
```

(ref:gss-trend-confidence) *Confidence in Scientific Community in the US*. The plot shows the variation of confidence in the Scientific Community in the US over time, for all answer options. The data for this plot is from the cumulative file of the US General Social Survey (GSS).

```{r gss-trend-confidence, fig.cap="(ref:gss-trend-confidence)"}
gss_summary_polarization |> 
  ggplot(aes(x = year, y = share, colour = confidence_factor)) +
  # Background: Other institutions with lower alpha
  geom_line(alpha = 0.2, size = 1) + 
 geom_point(size = 2) +
  # Apply Wes Anderson colors
  scale_color_manual(values = wes_palette("FantasticFox1", n = 4, type = "discrete")) +  
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +  
  scale_x_continuous(breaks = c(seq(min(gss_summary_polarization$year), max(gss_summary_polarization$year), by = 5), max(gss_summary_polarization$year))) +  

  labs(x = "Year", y = "Americans Confidence in Scientific Community", 
       caption = "Data source: GSS by NORC", color = "Level of Confidence") +
  theme_minimal() +  
  scale_alpha_identity()
```






