# Experiment 3 {#exp3}

## Materials

\FloatBarrier

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

We excluded all participants whose answer did not resemble "I pay attention".

## Descriptives

```{r exp3-descriptives}
# by discipline
exp3_wide %>% 
  group_by(discipline, treatment) %>% 
  summarise(across(c(n_correct, impressed, competence, trust), 
                   list(mean = ~mean(., na.rm=TRUE)), 
                   .names = "{col}_{fn}")) %>%
  mutate_if(is.numeric, ~round(.x, digits = 2)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = "Descriptives") %>%
  kable_styling(full_width = FALSE) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```

## Results

Fig. \@ref(fig:exp3-plot) illustrates the results.

(ref:exp3-plot) Overview of treatment effect on various outcomes

```{r exp3-plot, fig.cap="(ref:exp3-plot)"}
# ordre
plot_data <- exp3_long

plot_data$outcome <- factor(plot_data$outcome, levels = c("knowledge", "trust", "competence", "impressiveness"))
# by discipline
ggplot(plot_data, aes(x=outcome, y=value, fill = treatment)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = NULL, y = "Standardized likert scale rating", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) 


```

The results of the models from the hypotheses can be found in table \@ref(tab:exp3-hypotheses-table).

```{r exp3-hypotheses-table}
# H1 a
model_trust <- run_mixed_model(exp3_long, outcome = "trust")

# H1 b
model_impressed <- run_mixed_model(exp3_long, outcome = "impressiveness")

# H1 c
model_competence <- run_mixed_model(exp3_long, outcome = "competence")

modelsummary(list(
                  "Competence" = model_competence,
                  "Trust" = model_trust,
                  "Impressiveness" = model_impressed
                  ),
             stars = TRUE, 
             output = "kableExtra", 
             caption = "Models for interaction hypotheses") %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```

The treatment effects by outcome can be found in table \@ref(tab:exp3-treatment-effects).

```{r exp3-treatment-effects}
# Isolated Treatment effects
treatment_effect_knowledge <- lm(n_correct_std ~ treatment, exp3_wide) 

treatment_effect_trust <- lm(trust_std ~ treatment, exp3_wide)

treatment_effect_impressed <- lm(impressed_std ~ treatment, exp3_wide) 

treatment_effect_competence <- lm(competence_std ~ treatment, exp3_wide) 

modelsummary(list("Knowledge" = treatment_effect_knowledge,
                  "Trust" = treatment_effect_trust,
                  "Impressiveness" = treatment_effect_impressed,
                  "Competence" = treatment_effect_competence
), 
stars = TRUE, 
output = "kableExtra", 
caption = "Treatment effects by outcome") %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down")
```
