---
title: "Analysis Experiment 3"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
library(lsr)
library(effectsize)
library(readxl)
```

```{r, include=FALSE, message=FALSE}
data <- read_csv("/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Pilot 2/cleaned.csv")

data_wide <- read_xlsx("/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Pilot 2/wide_cleaned_with personal coding of recalls.xlsx")

# remove participants who did not complete phase 2 
#data_wide <- data_wide %>%
#  filter(!is.na(recall_t2))
```

**Note:** the coding of the quality of recall has been done with the help of chatgpt (prompt after giving it the two initial texts: "Here are summaries provided by participants in an experiment, who saw one of the two texts and then had to recall the information given in the text. Can you rate each summary depending on the quantity of information it provides, by giving one of three possible grades: 1, 2 and 3, from the least info to the more info.") This was done separately for summaries at t1 and summaries at t2. I changed the grade given by chatgpt one or two times when I considered it was not adapted.

# Descriptives

## Trust

```{r, echo=FALSE, message=FALSE, fig.show='hold'}
# means and confidence intervals of trust and competence by discipline at t0 and t1

## table
data_agg <- data %>% 
    pivot_longer(
      cols = c("competence", "trust"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
    group_by(discipline, time, outcome) %>% 
    summarize(mean = mean(score, na.rm = TRUE),
              CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(score, 0.95, na.rm = TRUE)[2]) %>% #upper bound 
  filter(!is.na(CI_up))

##figure
data %>% 
    pivot_longer(
      cols = c("competence", "trust"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
  ggplot(aes(x = outcome, y = score, fill = time)) + 
  geom_violin(alpha = 0.3, colour = "white") +
  geom_point(aes(x = outcome, y = mean, colour = time), data = data_agg, position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up, colour = time), 
                width = 0.2, 
                alpha = 1,
                data = data_agg,
                position = position_dodge(width = 0.9)) +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  facet_wrap(~discipline)

```

## Impressiveness

```{r}
# distribution of impressiveness ratings at t1
data %>%
  filter(!is.na(impressed)) %>%
  ggplot(aes(x = impressed, fill = discipline)) + 
  geom_bar(position = position_dodge(preserve = "single"), alpha = 0.8) +
  scale_fill_viridis_d(option = "magma", begin = 0.9, end = 0.95) +
  facet_wrap(~discipline)

```

## Recall Quality

```{r}
data_wide %>%
  pivot_longer(c("recall_t1_quality", "recall_t2_quality"), names_to = "time", values_to = "recall_quality") %>%
  mutate(time = case_when(time == "recall_t1_quality" ~ "t1", 
                          time == "recall_t2_quality" ~ "t2")) %>%
  ggplot(aes(x = recall_quality, fill = time)) + 
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_fill_viridis_d(option = "magma", begin = 0.9, end = 0.95) +
  facet_wrap(~main_discipline)
```

## Relation of recall quality to trust and impressiveness

```{r}
# trust, competence and impressiveness ratings depending on recall_quality at t2 

## summary table 
data_agg2 <- data_wide %>%
    pivot_longer(
      cols = c(starts_with("competence"), starts_with("trust"), starts_with("impressed")),  # Columns to pivot
      names_to = c("outcome", "time"),
      names_sep = "_",
      values_to = "score"                 # New column for values
    ) %>% 
    group_by(time, outcome, recall_t2_quality) %>% 
    summarize(mean = mean(score, na.rm = TRUE),
              CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(score, 0.95, na.rm = TRUE)[2],
              count = n()) %>% #upper bound 
  filter(!is.na(recall_t2_quality), !is.na(CI_up))

##figure
data_wide %>% 
    pivot_longer(
      cols = c(starts_with("competence"), starts_with("trust"), starts_with("impressed")),  # Columns to pivot
      names_to = c("outcome", "time"),
      names_sep = "_",
      values_to = "score"                 # New column for values
    ) %>% 
  filter(!is.na(score), !is.na(recall_t2_quality)) %>%
  ggplot(aes(x = recall_t2_quality, y = score, fill = time)) + 
  geom_point(aes(colour = time), alpha = 0.2, position = position_jitter(width = 0.1, height = 0.1)) +
  geom_point(aes(x = recall_t2_quality, y = mean, colour = time), data = data_agg2, position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up, colour = time), 
                width = 0.2, 
                alpha = 1,
                data = data_agg2,
                position = position_dodge(width = 0.9)) +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  facet_wrap(~outcome)

# difference in trust and competence between t0 and t1 depending on recall quality at t2
## summary table
data_agg3 <- data_wide %>%
    mutate(trust_diff = trust_t1 - trust_t0,
           competence_diff = competence_t1 - competence_t0)%>%
    pivot_longer(cols = c("trust_diff","competence_diff"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
    group_by(outcome, recall_t2_quality) %>% 
    summarize(mean = mean(score, na.rm = TRUE),
              CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(score, 0.95, na.rm = TRUE)[2],
              count = n()) %>% #upper bound 
  filter(!is.na(recall_t2_quality), !is.na(CI_up))

## figure 
data_wide %>%
    mutate(trust_diff = trust_t1 - trust_t0,
           competence_diff = competence_t1 - competence_t0)%>%
    pivot_longer(cols = c("trust_diff","competence_diff"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
  filter(!is.na(recall_t2_quality)) %>%
  ggplot(aes(x = recall_t2_quality, y = score)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0.15, height = 0.15)) +
  geom_point(aes(x = recall_t2_quality, y = mean), colour = "red", data = data_agg3) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up), 
                width = 0.2, 
                alpha = 1,
                colour= "red",
                data = data_agg3) +
  scale_x_continuous(n.breaks = 3) +
  facet_wrap(~outcome)
```

# Analysis

## Trust

```{r}
testtrust = t.test(data_wide$trust_t1, data_wide$trust_t0, paired = TRUE)
cdtrust = cohens_d(data_wide$trust_t1, data_wide$trust_t0, paired = TRUE)
testcomp = t.test(data_wide$competence_t1, data_wide$competence_t0, paired = TRUE)
cdcomp = cohens_d(data_wide$competence_t1, data_wide$competence_t0, paired = TRUE)

paste("trust t0-t1: ", "t(", round(testtrust$parameter, 3), ") = ",  round(testtrust$statistic,3), ", p = ", round(testtrust$p.value,3), ", Cohen's d = ", round(cdtrust$Cohens_d,2), " [", round(cdtrust$CI_low, 2), ",", round(cdtrust$CI_high, 2), "]",  sep = "")

paste("competence t0-t1: ", "t(", round(testcomp$parameter, 3), ") = ",  round(testcomp$statistic,3), ", p = ", round(testcomp$p.value,3), ", Cohen's d = ", round(cdcomp$Cohens_d,2), " [", round(cdcomp$CI_low, 2), ",", round(cdcomp$CI_high, 2), "]",  sep = "")
```
