---
title: "Analysis Experiment 4"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
library(lsr)
library(effectsize)
library(readxl)
```

```{r, include=FALSE, message=FALSE}
experiment_data <- read_xlsx("data/experiment_cleaned.xlsx")
exp_data_above_median <- read_xlsx("data/experiment_cleaned_above_median.xlsx")

validation_data <- read_xlsx("data/validation_cleaned.xlsx")
```

# Visualisations

## Knowledge scores

```{r, echo=FALSE, message=FALSE, fig.show='hold'}
## table with means and confidence intervals of knowledge scores by condition
data_agg_knowledge <- experiment_data %>% 
    group_by(condition) %>% 
    summarize(mean = mean(knowledge_score, na.rm = TRUE),
              CI_down = ciMean(knowledge_score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(knowledge_score, 0.95, na.rm = TRUE)[2]) #upper bound 

##figure
experiment_data %>% 
  ggplot(aes(x = knowledge_score, fill = condition)) + 
  geom_histogram(binwidth = 0.1, alpha = 0.8) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.2)) +
  scale_fill_viridis_d(option = "magma", begin = 0.5, end = 0.8) +
  facet_wrap(~condition)
```

## Competence, trust and impressiveness (above-median data)

```{r}
## table with means and confidence intervals
data_am_agg_abstrustcomp <- exp_data_above_median %>% 
    pivot_longer(
      cols = c("competence", "trust", "global_impressiveness"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
    group_by(condition, outcome) %>% 
    summarize(mean = mean(score, na.rm = TRUE),
              CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(score, 0.95, na.rm = TRUE)[2]) %>% #upper bound 
  filter(!is.na(CI_up))

##figure
exp_data_above_median %>% 
    pivot_longer(
      cols = c("competence", "trust", "global_impressiveness"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
  ggplot(aes(x = outcome, y = score, fill = outcome)) + 
  geom_violin(alpha = 0.3, colour = "white") +
  geom_point(aes(x = outcome, y = mean, colour = outcome), data = data_am_agg_abstrustcomp, position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up, colour = outcome), 
                width = 0.2, 
                alpha = 1,
                data = data_am_agg_abstrustcomp,
                position = position_dodge(width = 0.9)) +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.7) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.7) +
  facet_wrap(~condition)
```

## Change in competence and trust (above-median data)

```{r}
## table with means and confidence intervals
data_am_agg_trustcomp <- exp_data_above_median %>% 
  pivot_longer(c("change_competence", "change_trust"), names_to = "outcome", values_to = "score") %>%
  group_by(condition, outcome) %>% 
  summarize(mean = mean(score, na.rm = TRUE),
            CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
            CI_up = ciMean(score, 0.95, na.rm = TRUE)[2]) #upper bound 

## figure
exp_data_above_median %>% 
  pivot_longer(c("change_competence", "change_trust"), names_to = "outcome", values_to = "score") %>%
  ggplot(aes(x = score, colour = outcome, fill = outcome)) +
  geom_bar(alpha = 0.8, colour = "white", position = "dodge")  +
  scale_y_continuous(name = "count") + 
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  facet_wrap(~condition) 

```

## Impressive knowledge

```{r}

## table with means and confidence intervals
data_agg_impknow <- experiment_data %>% 
  group_by(condition) %>% 
  summarize(mean = mean(n_impressive, na.rm = TRUE),
            CI_down = ciMean(n_impressive, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
            CI_up = ciMean(n_impressive, 0.95, na.rm = TRUE)[2]) #upper bound 

## figure
experiment_data %>% 
  ggplot(aes(x = n_impressive, fill = condition)) +
  geom_bar(alpha = 0.8, colour = "white", position = "dodge")  +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  facet_wrap(~condition) 

```

## Impressive knowledge (above-median data)

```{r}
## table with means and confidence intervals
data_agg_impknow <- exp_data_above_median %>% 
  group_by(condition) %>% 
  summarize(mean = mean(n_impressive, na.rm = TRUE),
            CI_down = ciMean(n_impressive, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
            CI_up = ciMean(n_impressive, 0.95, na.rm = TRUE)[2]) #upper bound 

## figure
exp_data_above_median %>% 
  ggplot(aes(x = n_impressive, fill = condition)) +
  geom_bar(alpha = 0.8, colour = "white", position = "dodge")  +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  facet_wrap(~condition) 
```

## Forgetting scores (above-median data)

```{r}
## table with means and confidence intervals
data_am_agg_forget <- exp_data_above_median %>% 
  pivot_longer(c("forgetting_score", "impressive_forgetting_score"), names_to = "outcome", values_to = "score") %>%
  group_by(condition, outcome) %>% 
  summarize(mean = mean(score, na.rm = TRUE),
            CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
            CI_up = ciMean(score, 0.95, na.rm = TRUE)[2]) #upper bound 

## figure
exp_data_above_median %>% 
  pivot_longer(c("forgetting_score", "impressive_forgetting_score"), names_to = "outcome", values_to = "score") %>%
  ggplot(aes(x = outcome, y = score, colour = outcome, fill = outcome)) +
  geom_violin(alpha = 0.3, colour = "white", position = "dodge")  +
  geom_point(aes(x = outcome, y = mean, colour = outcome), data = data_am_agg_forget, position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up, colour = outcome), 
                width = 0.2, 
                alpha = 1,
                data = data_am_agg_forget,
                position = position_dodge(width = 0.9)) +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  facet_wrap(~condition)
```

## Validation study

```{r}
## table
data_val_agg <- validation_data %>% 
    pivot_longer(
      cols = c("competence", "trust", "impressiveness"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
    group_by(condition, outcome) %>% 
    summarize(mean = mean(score, na.rm = TRUE),
              CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(score, 0.95, na.rm = TRUE)[2]) %>% #upper bound 
  filter(!is.na(CI_up))

##figure
validation_data %>% 
    pivot_longer(
      cols = c("competence", "trust", "impressiveness"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
  ggplot(aes(x = outcome, y = score, fill = condition)) + 
  geom_violin(alpha = 0.3, colour = "white") +
  geom_point(aes(x = outcome, y = mean, colour = condition), data = data_val_agg, position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up, colour = condition), 
                width = 0.2, 
                alpha = 1,
                data = data_val_agg,
                position = position_dodge(width = 0.9)) +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.5) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.5)
```

# Analysis

```{r}
# initialize an empty results data frame
results <- data.frame(hypothesis = character(), test_type = character(), statistic = numeric(), p_value = numeric(), mean_difference = numeric(), stringsAsFactors = FALSE
)
```

## H1a and H1b (above-median data)

```{r}
# H1a: One-sample t-test for change in competence
if (shapiro.test(exp_data_above_median$change_competence)$p.value > 0.05) {
  h1a <- t.test(exp_data_above_median$change_competence, mu = 0)
  results <- rbind(results, data.frame(
    hypothesis = "H1a",
    test_type = "one sample t-test",
    statistic = h1a$statistic,
    p_value = h1a$p.value,
    mean_difference = h1a$estimate
  ), row.names = NULL)
} else {
  h1a <- wilcox.test(exp_data_above_median$change_competence, mu = 0, exact = FALSE)
  results <- rbind(results, data.frame(
    hypothesis = "H1a",
    test_type = "Wilcoxon",
    statistic = h1a$statistic,
    p_value = h1a$p.value,
    mean_difference = NA
  ), row.names = NULL)
}

# H1b: One-sample t-test for change in trust
if (shapiro.test(exp_data_above_median$change_trust)$p.value > 0.05) {
  h1b <- t.test(exp_data_above_median$change_trust, mu = 0)
  results <- rbind(results, data.frame(
    hypothesis = "H1b",
    test_type = "one sample t-test",
    statistic = h1b$statistic,
    p_value = h1b$p.value,
    mean_difference = h1b$estimate
  ), row.names = NULL)
} else {
  h1b <- wilcox.test(exp_data_above_median$change_trust, mu = 0, exact = FALSE)
  results <- rbind(results, data.frame(
    hypothesis = "H1b",
    test_type = "Wilcoxon",
    statistic = h1b$statistic,
    p_value = h1b$p.value,
    mean_difference = NA
  ), row.names = NULL)
}

```

## H2 (above-median data)

```{r}
# H2: Test forgetting score against zero
if (shapiro.test(exp_data_above_median$forgetting_score)$p.value > 0.05) {
  h2 <- t.test(exp_data_above_median$forgetting_score, mu = 0)
  results <- rbind(results, data.frame(
    hypothesis = "H2",
    test_type = "one sample t-test",
    statistic = h2$statistic,
    p_value = h2$p.value,
    mean_difference = h2$estimate
  ), row.names = NULL)
} else {
  h2 <- wilcox.test(exp_data_above_median$forgetting_score, mu = 0, exact = FALSE)
  results <- rbind(results, data.frame(
    hypothesis = "H2",
    test_type = "Wilcoxon",
    statistic = h2$statistic,
    p_value = h2$p.value,
    mean_difference = NA
  ), row.names = NULL)
}

```

## H3 (above-median data)

```{r}
# H3: One-sample test for impressive forgetting score against zero
if (shapiro.test(exp_data_above_median$impressive_forgetting_score)$p.value > 0.05) {
  h3 <- t.test(exp_data_above_median$impressive_forgetting_score, mu = 0)
  results <- rbind(results, data.frame(
    hypothesis = "H3",
    test_type = "one sample t-test",
    statistic = h3$statistic,
    p_value = h3$p.value,
    mean_difference = h3$estimate
  ), row.names = NULL)
} else {
  h3 <- wilcox.test(exp_data_above_median$impressive_forgetting_score, mu = 0, exact = FALSE)
  results <- rbind(results, data.frame(
    hypothesis = "H3",
    test_type = "Wilcoxon",
    statistic = h3$statistic,
    p_value = h3$p.value,
    mean_difference = NA
  ), row.names = NULL)
}

```

## H4a to H4c

```{r}
# H4a: Independent t-test for text impressiveness
h4a <- t.test(
  impressiveness ~ condition,
  data = validation_data
)
results <- rbind(results, data.frame(
  hypothesis = "H4a",
  test_type = "independent t-test",
  statistic = h4a$statistic,
  p_value = h4a$p.value,
  mean_difference = h4a$estimate[[1]] - h4a$estimate[[2]]
))

# H4b: Independent t-test for competence change
h4b <- t.test(
  competence ~ condition,
  data = validation_data
)
results <- rbind(results, data.frame(
  hypothesis = "H4b",
  test_type = "independent t-test",
  statistic = h4b$statistic,
  p_value = h4b$p.value,
  mean_difference = h4b$estimate[[1]] - h4b$estimate[[2]]
))

# H4c: Independent t-test for trust change
h4c <- t.test(
  trust ~ condition,
  data = validation_data
)
results <- rbind(results, data.frame(
  hypothesis = "H4c",
  test_type = "independent t-test",
  statistic = h4c$statistic,
  p_value = h4c$p.value,
  mean_difference = h4c$estimate[[1]] - h4c$estimate[[2]]
))
```

## Exploratory : trust and competence for below-median vs above-median participants

```{r}
# Calculate cut off values 
cutoff_archeo50 = quantile(experiment_data$knowledge_score[experiment_data$condition == "archeo"], 0.50)
cutoff_entom50 = quantile(experiment_data$knowledge_score[experiment_data$condition == "entom"], 0.50)

# Create column indicating whether above the cut-off
experiment_data <- experiment_data %>%
  mutate(above_median = case_when(condition == "archeo" ~ knowledge_score >= cutoff_archeo50,
                                  condition == "entom" ~ knowledge_score >= cutoff_entom50)) 

## table with means and confidence intervals
data_median_agg_abstrustcomp <- experiment_data %>% 
    pivot_longer(
      cols = c("competence", "trust", "global_impressiveness"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
    group_by(above_median, condition, outcome) %>% 
    summarize(mean = mean(score, na.rm = TRUE),
              CI_down = ciMean(score, 0.95, na.rm = TRUE)[1], #lower bound of 95% Confidence interval around the mean
              CI_up = ciMean(score, 0.95, na.rm = TRUE)[2]) %>% #upper bound 
  filter(!is.na(CI_up))

##figure
experiment_data %>% 
    pivot_longer(
      cols = c("competence", "trust", "global_impressiveness"),  # Columns to pivot
      names_to = c("outcome"),
      values_to = "score"                 # New column for values
    ) %>% 
  ggplot(aes(x = outcome, y = score, fill = above_median)) + 
  geom_violin(alpha = 0.3, colour = "white") +
  geom_point(aes(x = outcome, y = mean, colour = above_median), data = data_median_agg_abstrustcomp, position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(y = mean, ymin = CI_down, ymax = CI_up, colour = above_median), 
                width = 0.2, 
                alpha = 1,
                data = data_median_agg_abstrustcomp,
                position = position_dodge(width = 0.9)) +
  scale_fill_viridis_d(option = "magma", begin = 0.2, end = 0.7) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.7) +
  facet_wrap(~condition) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )
```
