---
title: "Coding reliability analysis"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
library(lsr)
library(effectsize)
library(readxl)
library("irr")
library(writexl)
```

# Make spreadsheet template for coders

```{r}
# set a seed for random sampling
set.seed(9823)
```


We begin by using the chat gpt coding tables generated in the `cleaning.Rmd` file. 

```{r}
archeo <- read_xlsx("data/archeo_gptcoding_with_justification.xlsx")
entom <- read_xlsx("data/entom_gptcoding_with_justification.xlsx")
```

We first remove all content.

```{r}
remove_chatgpt_codings <- function(df) {
  df <- df |> 
  mutate(across(-c(id, condition, text), ~ NA))
}

archeo <- remove_chatgpt_codings(archeo)
entom <- remove_chatgpt_codings(entom)
```

We then randomly pick 40 texts per condition, i.e. 80 in total

```{r}
sample <- function(df) {
  df <- df |> 
    sample_n(40)
}

archeo <- sample(archeo)
entom <- sample(entom)
```

Then we then export them as a compbineddata frame. 

```{r}
coder_data <- bind_rows(archeo, entom) 

write_csv(coder_data, ("data/template_human_coders.csv"))
```

# Import data

```{r, include=FALSE, message=FALSE}
coding1 <- read_xlsx("/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Coding participants' texts/coding_pilot2_Sophie.xlsx", skip = 1) %>%
  #add the coder ID to the column names (ex: for coder number 1, rating columns start wity c1_)
  rename_with(~ gsub("cX", "c1", .x))

coding2 <- read_xlsx("/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Coding participants' texts/coding_pilot2_Jan.xlsx", skip = 1) %>%
  rename_with(~ gsub("cX", "c2", .x))

codingGPT <- read_xlsx("/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Coding participants' texts/coding_pilot2_gpt.xlsx") %>%
  rename_with(~ gsub("cX", "cGPT", .x))

# append data 
coding2 <- coding2 %>%
  select(-participant_ID, -discipline, -text)
codingGPT <- codingGPT %>%
  select(-participant_ID, -discipline, -text, -text_corrected)
data = full_join(coding1, coding2, by = "text_ID") %>%
  full_join(codingGPT, by = "text_ID")
```

# Cleaning

```{r}
# compute proportion of knowledge items present in the text by dividing total knowledge scores by maximum possible score 
maxscore_archeo = ncol(coding1 %>% select(contains("archeo")))*2
maxscore_entom = ncol(coding1 %>% select(contains("entom")))*2
                         

data <- data %>%
  mutate(c1_totalp = case_when(discipline == "archeo" ~ c1_total/maxscore_archeo,
                               discipline == "entom" ~ c1_total/maxscore_entom), 
         c2_totalp = case_when(discipline == "archeo" ~ c2_total/maxscore_archeo,
                               discipline == "entom" ~ c2_total/maxscore_entom),
         cGPT_totalp = case_when(discipline == "archeo" ~ cGPT_total/maxscore_archeo,
                               discipline == "entom" ~ cGPT_total/maxscore_entom))

data_long <- data %>%
  pivot_longer(starts_with("c"), 
               names_to = c("coder", "knowledge_dim"), 
               names_sep = "_", 
               values_to = "rating")%>%
  filter(!is.na(rating))

```

# Check distribution of ratings

```{r}
data_long %>%
  filter(knowledge_dim == "totalp") %>%
  ggplot(aes(x = rating)) + 
  geom_histogram()
```

# ICC

#### Function to compute ICC

Code to compute ICC comes from ten Hove, D., Jorgensen, T. D., & van der Ark, L. A. (2022), "Updated Guidelines on Selecting an Intraclass Correlation Coefficient for Interrater Reliability, With Applications to Incomplete Observational Designs". Psychological Methods. Advance online publication. <https://dx.doi.org/10.1037/met0000516>). The code is available at <https://osf.io/8j26u/>.

**Note:** "if(all(rowSums(table(data$rater, by = data$subject)) == 1))" corrected as "if(all(rowSums(table(data[[raters]], by = data[[subjects]])) == 1))"

```{r}
## Updated Guidelines on Selecting an ICC for Interrater Reliability: 
## with Applications to Incomplete Observational Designs
## Author: Debby ten Hove

## Supplement 2: ICC functions 
##    - computeQkhat: Computes Q and khat from observational design
##    - computekhat: Computes khat from observational design
##    - estICCs: Estimates ICCs and variance components with MCMC and MLE4

## Required packages:
## - brms
## - rstan
## - lme4
## - merDeriv
## - car

## Install missing packages
required.packages <- c("brms", "rstan", "lme4", "merDeriv", "car")
to.install <- required.packages[!(required.packages %in% installed.packages()[,"Package"])]
if(length(to.install)) install.packages(to.install) 

############################################################################
## FUNCTIONS TO COMPUTE KHAT AND Q
############################################################################
## Function to compute khat and Q from obs. design
computeQkhat <- function(data, subjects = "subjects", raters = "raters"){
  k <- length(unique(data[[raters]]))
  share <- 0 # Proportion shared raters
  
  RR <- data[[raters]]
  SS <- data[[subjects]]
  tabRxS <- table(RR, SS)
  uSub <- ncol(tabRxS) # Number of unique subjects
  khat <- uSub/ sum(1/colSums(tabRxS)) # harmonic mean number of raters per subject
  for(i in 1:uSub){
    k_s <- colSums(tabRxS)[i]
    
    for(j in (1:uSub)[-i]){
      k_sprime <- colSums(tabRxS)[j]
      k_s.sprime <- sum(RR[SS == i] %in% RR[SS == j])
      share <- share + (k_s.sprime / (k_s*k_sprime))/(uSub * (uSub-1))
    }
  }
  Q <- round(1/khat - share, 3)
  names(Q) <- "Q"
  
  return(list(Q = Q, khat = khat))
}

## Function to compute only khat (saves time when Q is not needed)
computeKhat <- function(data, subjects = "subjects", raters = "raters"){
  k <- length(unique(data[[raters]]))
  share <- 0 # Proportion shared raters
  
  RR <- data[[raters]]
  SS <- data[[subjects]]
  tabRxS <- table(RR, SS)
  uSub <- ncol(tabRxS) # Number of unique subjects
  khat <- uSub/ sum(1/colSums(tabRxS)) # harmonic mean number of raters per subject
  
  return(khat)
}
############################################################################
## FUNCTION TO ESTIMATE ICCs
############################################################################
estICCs <- function(data, 
                    Y = "Y", 
                    subjects = "subjects", 
                    raters = "raters", 
                    level = .95,
                    k = NULL, 
                    khat = NULL, 
                    Q = NULL, 
                    estimator = "MLE") {
  
  ## Number of raters
  if(is.null(k)){
    k <- length(unique(data[[raters]]))
  }
  
  ## Check type of design
  # Balanced or unbalanced
  if(length(unique(rowSums(table(data[[subjects]], by = data[[raters]])))) == 1){
    balanced <- "TRUE"
  } else {
    balanced <- "FALSE"
  }
  # Complete or incomplete
  if(unique(colSums(table(data[[subjects]], by = data[[raters]])) == 
            length(unique(data[[subjects]])))){
    complete <- "TRUE" 
  } else {
    complete <- "FALSE"
  }
  # One-Way or Two-Way
  if(all(rowSums(table(data[[raters]], by = data[[subjects]])) == 1)){
    twoWay <- "FALSE"
  } else {
    twoWay <- "TRUE"
  }
  
  if(is.null(khat) | is.null(Q)){
    ## Decide on values for khat and q 
    if(balanced == T & complete == T){ 
      khat <- k
      Q <- 0 
    } else {
      if(balanced == T & complete == F){
        khat <- unique(rowSums(table(data[[subjects]], by = data[[raters]])))
        Q <- computeQkhat(data, subjects = subjects, raters = raters)$Q
      } else {
        if(balanced == F & complete == F){
          Qkhat <- computeQkhat(data, subjects = subjects, raters = raters)
          khat <- Qkhat$khat
          Q <- Qkhat$Q
        } else {
          if(twoWay == F){
            khat <- computeKhat(data, subjects = subjects, raters = raters)
            Q <- 1/k # But not needed, since sigmaR cannot be distinguished
          }
        }
      }
    }
  }
  
  ## ICC, sigma and SD names, for indexing and renaming
  ICCnames <- c("ICCa1", "ICCak", "ICCakhat",
                "ICCc1", "ICCck", "ICCqkhat")
  SDnames <- c("SD_s", "SD_r", "SD_sr") 
  sigmanames <- c("S_s", "S_r", "S_sr") 
  
  ## Variable names of output
  outnames <-  c(ICCnames, 
                 paste0(rep(ICCnames, each = 2), c(".l", ".u")), 
                 paste0(ICCnames, "_se"), 
                 SDnames, 
                 paste0(rep(SDnames, each = 2), c(".l", ".u")), 
                 paste0(SDnames, "_se"), 
                 sigmanames, 
                 paste0(rep(sigmanames, each = 2), c(".l", ".u")), 
                 paste0(sigmanames, "_se"),
                 "Q", "khat", "k", "time")
  
  if(estimator == "MCMC"){
  ## Estimate two-way model using BRMS
    modForm <- paste(Y, "~ 1 + (1|", subjects, ") + (1 |", raters, ")")
    ## Estimate model
    brmOUT <- brms::brm(as.formula(modForm),
                      data   = data, 
                      warmup = 500, 
                      iter   = 1000, 
                      chains = 3, 
                      inits  = "random",
                      cores  = 3)
  # IF non-converged: Return NA for everything 
  if(any( abs(brms::rhat(brmOUT)[2:4] - 1) > .1)){ 
    MCMC <- rep(NA, times = length(outnames))
    names(MCMC) <- outnames 
  } else {
    ## If converged: Give me results 
    ## Extract posterior distribution of SDs
    SDs <- rstan::extract(brmOUT$fit, c(paste0("sd_", subjects, "__Intercept"),
                                        paste0("sd_",   raters, "__Intercept"),
                                        "sigma"))
    # List SDs to later get MAPs
    names(SDs) <- c("SD_s", "SD_r", "SD_sr")
    
    ## Convert to variances 
    S_s <- SDs[["SD_s"]]^2
    S_r <- SDs[["SD_r"]]^2  
    S_sr <- SDs[["SD_sr"]]^2
    
    # List variances to later get MAPs
    sigmas <- list(S_s = S_s, S_r = S_r, S_sr = S_sr)
    
    ## Obtain ICCs
    ICCa1 <- as.numeric(S_s / (S_s + S_r + S_sr))
    ICCak <- as.numeric(S_s / (S_s + (S_r + S_sr)/k))
    ICCakhat <- as.numeric(S_s / (S_s + (S_r + S_sr)/khat))
    ICCc1 <- as.numeric(S_s / (S_s + S_sr))
    ICCck <- as.numeric(S_s / (S_s + S_sr/k))
    ICCqkhat <- as.numeric(S_s / (S_s + Q*S_r + S_sr/khat))
    
    
    ICCs <- list(ICCa1 = ICCa1, ICCak = ICCak, ICCakhat = ICCakhat,
                 ICCc1 = ICCc1, ICCck = ICCck, ICCqkhat = ICCqkhat)
    
    ## Confidence levels 
    ICC_cis <-  do.call(rbind, lapply(ICCs, quantile, probs = c((1-level)/2, level + (1-level)/2)))
    sigma_cis <- do.call(rbind, lapply(sigmas, quantile, probs = c((1-level)/2, level + (1-level)/2)))
    
    ## SEs (posterior SDs)
    ICC_ses <- unlist(lapply(ICCs, sd))
    sigma_ses <- unlist(lapply(sigmas, sd))

    ## Point estimates (last, to not overwrite sigmas, SDs and ICCs sooner)
    # function to estimate posterior modes
    Mode <- function(x) {
      d <- density(x)
      d$x[which.max(d$y)]
    }
    # MAPs 
    sigmas <- mapply(Mode, sigmas) # Variances
    ICCs <- mapply(Mode, ICCs) # ICCs 
    
    # combine results
    ICCs <- cbind(ICCs, ICC_cis, ICC_ses)
    colnames(ICCs) <- c("ICC", "lower", "upper", "se")
    sigmas <- cbind(sigmas, sigma_cis, sigma_ses)
    colnames(ICCs) <- c("variance", "lower", "upper", "se")
    out <- list(ICCs = ICCs, 
                sigmas = sigmas, 
                Q = Q, khat = khat, k = k)
    
    }
  }
  
  if(estimator == "MLE"){
  ### Estimate two-way model using lme4 (random-effects model)
  ## Define model
  modForm <- paste(Y, "~ 1 + (1|", subjects, ") + (1 |", raters, ")")
  ## Estimate model
  mod   <- lme4::lmer(as.formula(modForm), data = data)
  ## Check convergence
  checkConv <- function(mod) { 
    warn <- mod@optinfo$conv$lme4$messages
    !is.null(warn) && grepl('failed to converge', warn) 
  }
  if(checkConv(mod)){
    # If nonconverged: Return NAs for everything
    MLE4 <- rep(NA, times = length(outnames))
    names(MLE4) <- outnames
  } else {
    
    # If converged: Give results
    ## Extract variances
    S_s  <- lme4::VarCorr(mod)[[subjects]][1, 1]  
    S_r  <- lme4::VarCorr(mod)[[raters]][1, 1]
    S_sr  <- sigma(mod)^2 
    
    ## Compute ICC point estimates
    ICCa1 <- S_s / (S_s + S_r + S_sr)
    ICCak <- S_s / (S_s + (S_r + S_sr)/k)
    ICCakhat <- S_s / (S_s + (S_r + S_sr)/khat)
    ICCc1 <- S_s / (S_s + S_sr) 
    ICCck <- S_s / (S_s + (S_sr)/k) 
    ICCqkhat <- S_s / (S_s + Q*S_r + S_sr/khat)
    
    ## List all (and create SDs)
    sigmas <- c(S_s = S_s, S_r = S_r, S_sr = S_sr)
    ICCs <- c(ICCa1 = ICCa1, ICCak = ICCak, ICCakhat = ICCakhat,
              ICCc1 = ICCc1, ICCck = ICCck, ICCqkhat = ICCqkhat)
    
    ## Asymptotic vcov matrix of sigmas
    suppressWarnings(ACOV <- merDeriv::vcov.lmerMod(mod, full = TRUE))
    Sidx <- grep(pattern = subjects, colnames(ACOV), fixed = TRUE) 
    Ridx <- grep(pattern = raters, colnames(ACOV), fixed = TRUE)
    SRidx <- which(colnames(ACOV) == "residual")
    idx      <- c(   Sidx  ,  Ridx  ,  SRidx  )
    newNames <- c("subject", "rater", "interaction")
    VCOV <- ACOV[idx, idx]
    dimnames(VCOV) <- list(newNames, newNames)
    vars <- c(subject = S_s, rater = S_r, interaction = S_sr)
    
    ## CIs and SEs of ICCs using asymptotic vcov matrix
    ## All info of all ICCs in one list
    ICCdefs <- c("subject / (subject + rater + interaction)", 
                 "subject / (subject + (rater + interaction)/k)",
                 "subject / (subject + (rater + interaction)/khat)", 
                 "subject / (subject + interaction)",
                 "subject / (subject + interaction/k)",
                 "subject / (subject + Q*rater + interaction/khat)"
    )
    names(ICCdefs) <- ICCnames
    ICCs_dm <- do.call("rbind", lapply(ICCdefs, FUN = function(x){
      car::deltaMethod(vars, vcov. = VCOV, level = level,g. = x)
    }))
    
    ICC_ses <- ICCs_dm[,"SE"]
    names(ICC_ses) <- paste0(ICCnames, "_se")
    sigma_ses <- do.call("rbind", lapply(newNames, FUN = function(x){
      car::deltaMethod(vars, vcov. = VCOV, level = level,g. = x)
    }))$SE 
    names(sigma_ses) <- paste0(names(sigmas), "_se")
    
    ## Monte-Carlo CIs of variances and ICCs
    dimnames(VCOV) <- list(names(sigmas), names(sigmas))
    sigma_mcCIs <- semTools::monteCarloCI(expr = c(S_s = 'S_s', S_r = "S_r", S_sr = "S_sr"),
                                          coefs = sigmas, ACM = VCOV)
    ICC_mcCIs <- semTools::monteCarloCI(expr = c(ICCa1_ci = "S_s / (S_s + S_r + S_sr)", 
                                                 ICCak_ci = paste0("S_s / (S_s + (S_r + S_sr)/", k, ")"),
                                                 ICCakhat_ci = paste0("S_s / (S_s + (S_r + S_sr)/", khat, ")"), 
                                                 ICCc1_ci = "S_s / (S_s + S_sr)",
                                                 ICCck_ci = paste0("S_s / (S_s + S_sr/", k, ")"),
                                                 ICCqkhat_ci = paste0("S_s / (S_s + ", Q, "*S_r + S_sr/", khat, ")")),
                                        coefs = sigmas, ACM = VCOV)
    
    ## Results in matrices
    ICCs <- cbind(ICC_mcCIs, ICC_ses)
    sigmas <- cbind(sigma_mcCIs, sigma_ses)
      dimnames(ICCs) <- list(outnames[1:6], c("ICC", "lower", "upper", "se"))
      colnames(sigmas) <- c("variance", "lower", "upper", "se")
      
    out <- list(ICCs = ICCs, sigmas = sigmas, Q = Q, khat = khat, k = k)
    }
  }
  
  ## return results
  return(out)
}
```

#### Compute ICC

```{r}
ICC = data_long %>%
  filter(knowledge_dim == "totalp") %>%
  estICCs(Y = "rating", subjects = "text_ID", raters = "coder", estimator = "MLE")

#extracting ICC for agreement between single ratings of random raters 
ICC$ICCs["ICCa1",]
```

# Percentage of agreement

```{r}

# coder 1 vs chatgpt 
agr_c1_GPT = c()
for(dim in knowledge_dim){
  column1 = paste("c1_", dim, sep = "")
  column2 = paste("cGPT_", dim, sep = "")
  d = data %>% select(column1,column2)
  agr = (sum(data[column1] == data[column2], na.rm = TRUE)/sum(!is.na(data[column1] == data[column2])))*100
  agr_c1_GPT = append(agr_c1_GPT, agr)
}

# coder 2 vs chatgpt 
agr_c2_GPT = c()
for(dim in knowledge_dim){
  column1 = paste("c2_", dim, sep = "")
  column2 = paste("cGPT_", dim, sep = "")
  d = data %>% select(column1,column2)
  agr = (sum(data[column1] == data[column2], na.rm = TRUE)/sum(!is.na(data[column1] == data[column2])))*100
  agr_c2_GPT = append(agr_c2_GPT, agr)
}

#coder 1 vs coder 2 
agr_c1_c2 = c()
for(dim in knowledge_dim){
  column1 = paste("c1_", dim, sep = "")
  column2 = paste("c2_", dim, sep = "")
  d = data %>% select(column1,column2)
  agr = (sum(data[column1] == data[column2], na.rm = TRUE)/sum(!is.na(data[column1] == data[column2])))*100
  agr_c1_c2 = append(agr_c1_c2, agr)
}
  
#create dataset
agr = data.frame(knowledge_dim, agr_c1_GPT, agr_c2_GPT, agr_c1_c2)
```

# Export data

```{r}
# wide dataset with all codings
write_xlsx(data, "/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Coding participants' texts/coding_combined.xlsx")

#long dataset with all codings
write_xlsx(data_long, "/Users/Sophie/Documents/Travail/Assistante de recherche - 2024/Confiance dans la science/Coding participants' texts/coding_combined_long.xlsx")

```
